---
title: "Untitled"
author: "irem"
date: '2022-08-31'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(cluster)
library(factoextra)
library(caret)
library(mice)
library(readxl)
```


```{r}
# importing data

wholesale <- read_excel("C:/Users/irem/Desktop/wholesale.xlsx")
```

First, let's get to know the variables:

*FRESH:* **Annual expenditure on fresh produce**
*MILK:* **Annual milk expenditure**
*GROCERY:* **Grocery products annual expenditure**
*FROZEN:* **Frozen products annual expenditure**
*DETERGENTS_PAPER:* **Detergent and paper annual expenditure**
*DELICATESSEN:* **Delicatessen annual expenditure**
*CHANNEL:* **(Hotel/Restoran/Cafe) or Retail channel (Nominal)**
*REGION:* **Customer area Lisnon, Oporto or other (Nominal)**


I scale units to get rid of outliers.

```{r}
scaleModel <- preProcess(wholesale, method = c("center","scale"))
modelData <- predict(scaleModel, wholesale)
```

I'm checking to see if there are any missing observations.

```{r}
md.pattern(modelData) #missing value does not exist.
```


```{r}
#Creating model
clusterModel <- kmeans(modelData, centers = 4, 
                iter.max = 15, nstart = 15)
clusterModel

fittedCluster <- fitted(clusterModel)

clusterModel$cluster
```

I'm doing a reverse scale here.

```{r}

reversedData <- modelData %>% select(one_of(scaleModel$mean %>% names)) %>%
                  map2_df(scaleModel$std, function(sd, var){var*sd})  %>%
                  map2_df(scaleModel$mean, function(mu, var){var+mu})

scaleModel$mean 
reversedData$cluster <- clusterModel$cluster


reversedData %>% group_by(cluster) %>% summarise_all(mean)
```



```{r}
boxplot(Fresh ~ cluster, data = reversedData,
        col="#80d7cc",border="pink")
```

The range in cluster 1 and the range in cluster 4 are almost the same. The median of the 3rd cluster and the 4th cluster is similar.

```{r}
fviz_cluster(clusterModel, data = modelData,
             ggtheme = theme_minimal()
             )
```


## Optimum "k" 

### k Calculation with Elbow Method

```{r}

clusterModel$withinss #cost function
clusterModel$tot.withinss 

wss <- sapply(2:10, FUN = function(x){kmeans(modelData, centers = x, 
                        nstart = 10, iter.max = 15)$tot.withinss})

wss

plot(2:10, wss, type = "b")

fviz_nbclust(modelData, kmeans, method = "wss")
```

### k C alculation with Silhouette Method

```{r}

silhouette(clusterModel$cluster, dist(modelData))

# The model with silhouette score max will be the best k model.

silScore <- function(x){
             model <- kmeans(modelData, centers = x , nstart = 10, iter.max = 15)
             sil <- silhouette(model$cluster, dist(modelData))[,3]
             score <- mean(sil)
             return(score)
             }

scores <- sapply(2:10, FUN = silScore)

scores

plot(2:10, scores, type = "b") 

```

Max score is best so "2" 

Let's sort.

```{r}
fviz_nbclust(modelData, kmeans, method = "silhouette")
```

According to the code above, the optimal 3 point came out because we gave nstart a random number.


```{r}

clusterModelK2 <- kmeans(modelData, centers = 2, nstart = 50, iter.max = 20)
clusterModelK3 <- kmeans(modelData, centers = 3, nstart = 50, iter.max = 20)

fviz_cluster(clusterModelK2, modelData)
```

In the code above there are outliers in the blue region at the extremes, the density is at the top.

```{r}
fviz_cluster(clusterModelK3, modelData)
```

In the above code, it created a cluster of outliers on the left.

```{r}

reversedData$clusterK2 <- clusterModelK2$cluster
reversedData$clusterK3 <- clusterModelK3$cluster

reversedData %>% group_by(clusterK2) %>% summarise_all(mean) 
```

More milk intake in cluster 2.

```{r}
reversedData %>% group_by(clusterK3) %>% summarise_all(mean)
```

When we look at the region, we see that there is no distinction according to this. Meat intake is higher in cluster 3.

*1 and 3 restaurants 2nd cluster hotel.*

As a result, we can segment customers according to this cluster and organize campaigns accordingly. The optimum number of clusters is 2 and 3. Whichever is more meaningful for the company, that is, profitable, we can use that cluster as customer segmentation.